---
title: "Tutorial R- data types"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# HI :)
Today I will tell you about data types in R. I don't know how much you learned so far, so if all of this is boring we will not waste time in person!:))


Your first hello world!
```{r}
print("Hello world!")
```
 
 You are a programmer now!:)
 
# Data types in R vs SPSS
 
 In SPSS, you (probably) mostly used: 
 
+ Numeric 
+ Comma / dot 
+ String
 
# Numeric and comma variables

```{r}
comma <- 1.2
class(comma) #function class() checks the class of the object
typeof(comma)

numeric <- 135
class(numeric)
typeof(numeric)
```
So usually R will see both as numeric and will probably assume it might have a decimal point. There is a specific integer type, to store numbers without decimal place, but for now it is not important. 

## Strings
 or in polish - *stringi* :) 

```{r}
string <- "Hello world!"
string2 <- "123" #If you put anything in quotation marks, it will be a string, not a number
string3 <- '1.234' #The same if you put it in single quotation marks
class(string)
class(string2)
class(string3)
```

As you can see, it does not matter if you use "" or ' ' . 
**Untill it does!**


```{r error = TRUE}
error <- "he said "stop!", everyone freezed"
```
As you can see, if your sentence has "" inside, it destroys your string and causes an error. 

```{r}
sentence <- "he said 'stop!', everyone freezed"
print(sentence)
```

So you need to use the other quotation mark outside, which is not used inside the sentence.

You can change the types. Remember our strings? 
```{r}
print(string2)
class(string2)

string2 <- as.numeric(string2)
print(string2)
class(string2)

print(string3)
class(string3)

string3 <- as.numeric(string3)
print(string3)
class(string3)

```
They are numbers now! **You are both a programmer and an alchemist now!**

There are also logical valriables.
*New data type unlocked*

```{r}
log <- TRUE
class(log)
```
Must be in caps lock tho. 

```{r error = TRUE}
log2 <- true
```

As you can see, R is case-sensitive. Which means that Variable and variable are different objects, the same goes for TRUE and true. SPSS was not case-sensitive, so Variable, variable, and VARIABLE are the same object.

However, R is not indent-sensitive. SPSS is not either, but some languages (e.g. Python) are. In other words, R does not care about white space in your code such as spaces and tabs. *Others reading your code might!*

```{r}
        print(
          
"I dont care!!!"
        )

  print(
              "I love it!!!"
   )
```

as a *fun fact* - R is indexing from 1. This might seem obvious to you, but most programming languages index from 0, not 1. 

```{r}
#R code
my_vector <- c("apple", "banana", "cherry")
print(my_vector[1]) 
print(my_vector[2])  

```

In python, if you point to 1, it not an apple!

```{python}

#Below code is written in python!

my_list = ["apple", "banana", "cherry"]
print(my_list[1])  #It would be a banana!
print(my_list[0])  #Apple is 0!

```

