---
title: "2. Data - cd"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

So now you know some basic data types in R. This lesson will cover tables, data frames and some basic operations.

So usually you do not have just one variable, rather a full dataset with multiple variables and data types. 

In SPSS you just had your dataset in one spreadsheet. Here, you will be able to have
```{r}


```

Importantly, if you make a copy of an object by assigning it to another object, they are seperate entities.

```{r}

original_vector <- c(1, 2, 3)

copied_vector <- original_vector

copied_vector <- c(copied_vector, 4)

print(paste("Original Vector:", toString(original_vector)))
print(paste("Copied Vector (modified):", toString(copied_vector)))

```

As you can see, the modification of our original vector did not influence the copied vector. **This is handy, because if you make a copy of your data set, you an perform operations on a copy, leaving the raw data untouched.** This could not be as straightforward in every language, for example in Python. 

```{python}

# Code written in python

original_list = [1, 2, 3]

copy_list = original_list

copy_list.append(4)

print("Original List:", original_list)
print("Copy List (modified):", copy_list)

```

Here as you can see operation on one variable changed them both. Fortunately, it is not our problem for now!


